### 2023-12-06 scripts/manuscript.qmd
recategorizing variables of interests

---
title: "The Social Conditioning of Fear"
author:
  - Alejandra Pedraza
  - W. Jamie Yang
abstract: "TBW" # this is where abstract should be pasted 
format: html
toc: true
toc-location: left
toc-title: Table of Contents
highlight-style: pygments
editor: visual
execute:
  eval: true # change to false before final submission
  echo: true
  message: false
  warning: false
fig-width: 6.5
fig-height: 8
code-fold: true
code-summary: "View Code"
embed-resources: true
bibliography: references.bib
link-citations: true
---

# Introduction

I'm leaving myself a note here to figure out how to do word count in a Quarto document. This will allow me to directly write in the qmd file and save the trouble of writing elsewhere like a word document and pasting things here.

# Methods

This study draws data from the General Social Survey (GSS). The GSS is a nationally representative interview-based survey that has collected longitudinal data since 1972. The GSS contains questions on demographics, behaviors, attitudes, civil liberties, crime and violence, intergroup tolerance, morality, national spending priorities, psychological well-being, social mobility, stress, and traumatic events. Our study is limited to data collected during the 2022 wave and we draw primarily on demographic and well-being variables.

Our dependent variable is coded as **fear** by the GSS. Respondents were asked, "*Is there any area right around here--that is, within a mile--where you would be afraid to walk alone at night?*" During the 2022 wave, there were 3,544 answers coded as follows: 878 = yes, 1440 = no, and 1226 = NA. Notably, the GSS codes missing values with multiple indicators--".i: Inapplicable," ".n: No answer," ".d: Do not Know/Cannot Choose," and ".s: Skipped on Web." For the purposes of our preliminary study, we coded all missing data as NA. Furthermore, we applied fixed removal of NAs for our dependent variable and we ran mice imputations for all other variables to create a more robust dataset that is necessary for further modeling.

Our control and independent variables are representative of two different dimensions we sought to capture. The first dimension is individual-level or social identity characteristics:

-   **born**: Were you born in this country?

-   **age**: Respondent's age in numeric values.

-   **sexnow1**: Do you describe yourself as male, female, or transgender?

-   **sexornt**: Which of the following best describes you?

-   **racerank1**: If you had to choose, which of these races do you identify most with?

-   **hispanic**: Are you Spanish, Hispanic, or Latino/Latina? IF YES: Which group are you from?

-   **degree**: Respondent's degree.

-   **uscitzn**: Are you a US citizen?

-   **dwelown**: (Do you/Does your family) own your (home/apartment), pay rent, or what?

Informed by theory, race is an important co-variate for our models as racialization serves to structure the daily lives of individuals based on where they fall on the Black-white spectrum, their opportunities, and their sense of belonging [@bonilla-silva2022; @omi2015]. First, we sought to create a combined race variable from the variables **race** and **hispanic**. After looking at the data, we saw that **race** was coded as "Black, White, or Other." This does not provide the control we need as it omits those in the "racial middle" that are more likely to identify with a panethnic label and reject the Black-white binary [@mora2014; @obrien2008; @rodriguez2000]. Instead, we opted to use the **racerank1** variable, because participants were able to select from 16 different options: White, Black or African-American, Asian Indian, Chinese, Filipino, Japanese, Korean, Vietnamese, Other, Asian, Guamanian or Chamorro, Samoan, Other Pacific Islander, Some other race, and Hispanic. We factored and recoded these to White, Black, Hispanic, Asian, NHPI (Native Hawaiian or Pacific Islander), AIAN (American Indian, Alaska Native), and Other. We understand that these categories are problematic and contentious [@haney-lópez2006; @dowling2014], however they are the standard with quantitative data.

Similarly, gender was a contentious variable to account for. Although the GSS has several questions aimed at capturing respondent's gender, we found them to be lacking in breadth. For example, the GSS codes **sex** within a Male/Female binary based solely on interviewer discretion. Ultimately, we opted to use **sexnow1** since it was the variable that most closely captured self-described gender. However, it remains problematic as respondents were only provided four choice: Male, Female, Transgender, and None of these. Out of 2306 cases, only 10 were coded as Transgender and 9 as None of these. Therefore, this is by no means a representative sample on gender minority population. The fact that the GSS data is primarily collected through in-person field interviews increases the possibility that minority-identifying individuals may not disclose their accurate identity information to avoid being associated with negative stereotypes.

Another important variable we sought to control for was citizenship status, since citizenship status and the associated stigma of illegality and threat of deportation have been found to increase fear in daily life [@golash-boza2015; @kline2019]. However, GSS data is not conducive to controlling for this variable due to the large number of NAs. Out of 2318 cases, 2164 were coded as NAs. In future work, we will pursue data options that allow the analysis ton include citizenship status.

The second dimension we sought to capture was community-level characteristics that would potentially affect one's sense of safety. For this dimension, we drew on the following variables:

-   **attend**: How often do you attend religious services?

-   **raclive**: Are there any ("whites" for black respondents, "blacks" for non-black respondents) living in this neighborhood now?

These variables seek to account for respondent's social embeddedness and networks of support. Specifically, **raclive** is an important variable to consider as it accounts for neighborhood composition, in the absence of a neighborhood dissimilarity index. While this variable only looks at Black-white racial composition, it provides a good benchmark to examine broader racial antagonisms. Research has documented the associated fears white experience when Blacks or people of other racial backgrounds move into their neighborhood [@lacayo2017; @massey2020].

The third dimension we considered were individual-level health-related and  attitudinal variables. Good mental and physical health status is positively correlated with individuals' evaluation of their life conditions and affects how likely individuals are to report positive emotions about their life [@dwyer2017]. Additionally, social capital--such as levels of interpersonal trust and norms of reciprocity and mutual aid--is believed to play an important role in explaining variations in health status across geographic localities [@kawachi1999]. Therefore, we included the following variables that capture respondents' self-reported health status and assessment of interpersonal social climate:

-   **happy**: Taken all together, how would you say things are these days--would you say that you are very happy, pretty happy, or not too happy?

-   **health**: Would you say your own health, in general, is excellent, good, fair, or poor?

-   **helpful**: Would you say that most of the time people try to be helpful, or that they are mostly just looking out for themselves?

-   **fair**: Do you think most people would try to take advantage of you if they got a chance, or would they try to be fair?

-   **trust**: Generally speaking, would you say that most people can be trusted or that you can't be too careful in dealing with people?

-   **partnrs5**: since February/March 2015, and including the past 12 months, how many sex partners have you had in that five year period?

## Data Script

```{r}
#| label: load-packages
#| include: false

library(tidyverse)
library(dplyr)
library(tidyr)
library(here)
library(rio)
library(janitor)
library(knitr)
library(kableExtra)
library(mice)
library(forcats)
library(ggplot2)
library(RColorBrewer)
library(spelling)
```

```{r}
#| label: load-data

gss <- import(here("data", "GSS.xlsx")) %>%
  as_tibble() %>% 
  clean_names()

```

```{r}
#| label: filtering-2022-data

gss2022 <- gss %>% 
  filter(year == "2022")

```

```{r}
#| label: select-variables

data <- gss2022 %>% 
    select(age, sex, sexnow1, sexornt, degree, attend, raclive, happy, health, helpful, fair, trust, fear, partnrs5, racerank1, uscitzn, dwelown)

```

```{r}
#| label: coding-missing-data-as-NAs

missing_vec <- c(".i:  Inapplicable", ".n:  No answer", ".d:  Do not Know/Cannot Choose", ".s:  Skipped on Web")

tidydata <- data %>% 
  mutate(across(c(age, sex, sexnow1, sexornt, degree, attend, raclive, happy, health, helpful, fair, trust, fear, partnrs5, racerank1, uscitzn, dwelown), 
                ~ifelse(. %in% missing_vec, NA_character_, .)))

```

```{r}
#| label: factoring-recoding-variables

tidydata <- tidydata %>% 
  filter(!is.na(fear)) %>% 
  mutate(age = as.numeric(age)) %>% 
  mutate(raclive = as.factor(raclive)) %>% 
  mutate(sexnow1 = as.factor(sexnow1)) %>% 
  mutate(fear = factor(case_when(fear == "NO" ~ "no",
                                 fear == "YES" ~ "yes"))) %>% 
  mutate(racecombo = 
           factor(case_when(racerank1 == "White" ~ "White",
                     racerank1 == "Black or African-American" ~ "Black",
                     racerank1 == "Hispanic" ~ "Hispanic",
                     racerank1 == "Asian Indian" | racerank1 ==  "Chinese" |
                       racerank1 ==  "Filipino" | racerank1 ==  "Japanese" | 
                       racerank1 ==  "Korean" |  racerank1 ==  "Vietnamese" |
                       racerank1 ==  "Other Asian" ~ "Asian",
                     racerank1 == "American Indian or Alaska Native" ~ "AIAN",
                     racerank1 == "Guamanian or Chamorro" | racerank1 == "Samoan" |
                       racerank1 == "Other Pacific Islander" ~ "NHPI",
                      racerank1 == "Some other race" ~ "Other"))) %>% 
  mutate(racecombo = fct_relevel(racecombo, 
              "White",
              "Black", 
              "Asian",
              "Hispanic",
              "AIAN",
              "NHPI",
              "Other")) %>%   
  mutate(partnrs5 = 
           factor(case_when(partnrs5 == "NO PARTNERS" ~ "0",
                            partnrs5 == "1 PARTNER" | partnrs5 == "2 PARTNERS" |
                              partnrs5 == "3 PARTNERS" | 
                              partnrs5 == "4 PARTNERS" ~ "Less than 5",
                            partnrs5 == "5-10 PARTNERS" ~ "5-10",
                            partnrs5 == "11-20 PARTNERS" ~ "11-20",
                            partnrs5 == "21-100 PARTNERS" ~ "21-100",
                            partnrs5 == "MORE THAN 100 PARTNERS" ~ "100 or more",
                            .default = NA))) %>%
  mutate(partnrs5 = fct_relevel(partnrs5,
                                "0",
                                "Less than 5",
                                "5-10",
                                "11-20",
                                "21-100",
                                "100 or more")) %>%
  mutate(uscitzn = 
           factor(case_when(uscitzn == "A u.s. citizen" | uscitzn == "A u.s. citizen born in puerto rico, the u.s. virgin islands, or the northern marianas islands (if volunteered)" |
                       uscitzn == "Born outside of the u.s. to parents who were u.s. citizens at that time (if volunteered)" ~ "Yes",
                     uscitzn == "Not a u.s. citizen" ~ "No",
                     .default = NA))) %>% 
   mutate(dwelown =
           factor(case_when(dwelown == "Own or is buying" ~ "Owner",
                            dwelown == "Pays rent" ~ "Renter",
                            dwelown == "Other" ~ "Other"))) %>% 
  mutate(degree = fct_relevel(degree, 
                              "Less than high school",
                              "High school",
                              "Associate/junior college",
                              "Bachelor's",
                              "Graduate")) %>% 
  mutate(attend = fct_relevel(attend, 
                              "Never",
                              "Less than once a year",
                              "About once or twice a year",
                              "2-3 times a month",
                              "Nearly every week",
                              "Every week",
                              "Several times a week")) %>% 
  mutate(happy = fct_relevel(happy, 
                             "Not too happy",
                             "Pretty happy",
                             "Very happy",)) %>% 
  mutate(health = fct_relevel(health, 
                              "Poor",
                              "Fair",
                              "Good",
                              "Excellent")) %>% 
  mutate(helpful = fct_relevel(helpful, 
                               "Looking out for themselves",
                               "Try to be helpful",
                               "Depends")) %>% 
  mutate(fair = fct_relevel(fair, 
                            "Would take advantage of you",
                            "Would try to be fair",
                            "Depends")) %>% 
  mutate(trust = fct_relevel(trust, 
                             "Can't be too careful",
                             "Can trust",
                             "Depends"))

```

```{r}
#| label: subset-data

gss_subset <- tidydata %>% 
  select(age, sexnow1, racecombo, uscitzn, degree, attend, raclive, happy, health, partnrs5, fear) 
```

```{r}
#| label: calculate-MeanFear

overall_fear <- gss_subset %>% 
  mutate(fear = case_when(fear == "no" ~ "0",
                                 fear == "yes" ~ "1")) %>% 
  select(fear)
mean_fear <- mean(as.numeric(overall_fear$fear, na.rm=TRUE))

```

```{r}
#| label: mice-imputation
#| eval: false

complete5 <- mice(gss_subset, 
                 m = 5, 
                 seed = 25)

summary(complete(complete5, 1))

temp5 <- complete(complete5, action = "long", include = TRUE)

tidydata_complete <- as.mids(temp5)

save(tidydata_complete, file = "data/tidydata_complete.RData")

```

# Results

In the absence of running models, in this section we will discuss descriptive statistics through a series of tables and plots. We seek to show how the experience of fear differed across different measures of individual-level and community-level characteristics. To begin with, across our sample of 2318 cases, mean fear was calculated at `mean_fear`.

First, we examined how fear is experienced across race and gender. As @tbl-FearRaceSex shows, Transgender and Female respondents had higher mean fear rate than respondents that identified as Male. Particularly, respondents that were Transgender expressed greater fear, but this may be indicative of sample representation. For example, only one respondent identified as both Hispanic and Transgender, and since they responded affirmative to the question, this finding is heavily skewed. When looking by race, whites were the group with the lowest mean fear and Asians were the group with the highest.

```{r}
#| label: tbl-FearRaceSex
#| tbl-cap: Mean fear by race and gender calculated from respondents affirmative answer to GSS question, Is there any area right around here--that is, within a mile--where you would be afraid to walk alone at night?

gss_subset %>% 
  filter(!is.na(racecombo)) %>% 
  filter(racecombo != "Other") %>% 
  filter(!is.na(sexnow1)) %>% 
  filter(sexnow1 != "None of these") %>% 
  pivot_longer(
    cols = fear,
    names_to = "fear",
    values_to = "answer") %>% 
  select(racecombo, answer, sexnow1) %>% 
  mutate(fear = as.numeric(case_when(answer == "no" ~ "0",
                                 answer == "yes" ~ "1"))) %>% 
  select(racecombo, fear, sexnow1) %>% 
  group_by(racecombo, sexnow1) %>% 
  summarize(mean_fear = mean(fear)) %>% 
  pivot_wider(
    names_from = sexnow1,
    values_from = mean_fear
  ) %>% 
  kable(caption = "Social Fear by Race and Gender",
        col.names = c("Race", "Female", "Male", "Transgender"),
        digits = 3) %>% 
  kableExtra::row_spec(0:6, background = "white") 
```

@fig-FearRaceGender does a good job at plotting this relationship.

```{r}
#| label: fig-FearRaceGender
#| fig-cap: Bar graph showing the percentage of respondants, by race and gender, that responded affirmative to GSS question, Is there any area right around here--that is, within a mile--where you would be afraid to walk alone at night?

gss_subset %>% 
  filter(!is.na(racecombo)) %>% 
  filter(racecombo != "Other") %>% 
  filter(!is.na(sexnow1)) %>% 
  filter(sexnow1 != "None of these") %>% 
  group_by(racecombo, sexnow1) %>% 
  summarize(fear = sum(fear == "yes"), 
            n = n(),
            prop = fear/n) %>% 
  ggplot(aes(x = racecombo, y = prop)) +
  geom_col(aes(fill = sexnow1), position = "dodge") + 
  scale_fill_brewer(palette = "Purples") +
  scale_y_continuous(labels = scales::percent) +
  labs(title = "Social Fear by Race and Gender",
       subtitle = "Is social fear influenced by respondent's race and gender?",
       x = "Race",
       y = "Percentage of Respondents who Expressed Fear") +
  theme_bw() +
  guides(fill = guide_legend(title = "Gender"))

```

Moving on to measures of social embeddedness, we examined fear across degrees of religious participation, degree of happiness and health, residential diversity, and sexual activity. Regarding religious participation, we did not discern any clear pattern, as @fig-FearReligiousPart displays below.

```{r}
#| label: fig-FearReligiousPart
#| fig-cap: Bar graph showing the percentage of respondants, by degree of participation in religious activities, that responded affirmative to GSS question, Is there any area right around here--that is, within a mile--where you would be afraid to walk alone at night?

gss_subset %>% 
  filter(!is.na(attend)) %>% 
  group_by(attend) %>% 
  summarize(fear = sum(fear == "yes"), 
            n = n(),
            prop = fear/n) %>% 
  ggplot(aes(x = attend, y = prop, fill = attend)) +
  geom_col() + 
  scale_fill_brewer(palette = "Blues") +
  scale_y_continuous(labels = scales::percent) +
  labs(title = "Social Fear by Relgious Participation",
       subtitle = "Is social fear influenced by religious participation?",
       x = "Religious Participation",
       y = "Percentage of Respondents who Expressed Fear") +
  theme_bw() +
  guides(fill = "none") +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))

```

Next, looking at measures of Happiness and Health, we found that generally as happiness and health increase, social fear decreases. This is shown in @fig-FearHappiness and @fig-FearHealth.

```{r}
#| label: fig-FearHappiness
#| fig-cap: Bar graph showing the percentage of respondants, by happiness levels, that responded affirmative to GSS question, Is there any area right around here--that is, within a mile--where you would be afraid to walk alone at night?

gss_subset %>% 
  filter(!is.na(happy)) %>% 
  group_by(happy) %>% 
  summarize(fear = sum(fear == "yes"), 
            n = n(),
            prop = fear/n) %>% 
  ggplot(aes(x = happy, y = prop, fill = happy)) +
  geom_col() + 
  scale_fill_brewer(palette = "PuBuGn") +
  scale_y_continuous(labels = scales::percent) +
  labs(title = "Social Fear by Happiness Level",
       subtitle = "Is social fear influenced by level of happiness?", 
       x = "Happiness Level",
       y = "Percentage of Respondents who Expressed Fear") +
  theme_bw() +
  guides(fill = "none") +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))
```

```{r}
#| label: fig-FearHealth
#| fig-cap: Bar graph showing the percentage of respondants, by health levels, that responded affirmative to GSS question, Is there any area right around here--that is, within a mile--where you would be afraid to walk alone at night?

gss_subset %>% 
  filter(!is.na(health)) %>% 
  group_by(health) %>% 
  summarize(fear = sum(fear == "yes"), 
            n = n(),
            prop = fear/n) %>% 
  ggplot(aes(x = health, y = prop, fill = health)) +
  geom_col() + 
  scale_fill_brewer(palette = "PuRd") +
  scale_y_continuous(labels = scales::percent) +
  labs(title = "Social Fear by Health Level",
       subtitle = "Is social fear influenced by health?", 
       x = "Health Level",
       y = "Percentage of Respondents who Expressed Fear") +
  theme_bw() +
  guides(fill = "none") +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))
```

Next, looking at residential composition and race combined displayed in @fig-FearNeighborhood, we do notice a pattern where respondents living in residentially more diverse areas were more likely to respond affirmative to the question of social fear.

```{r}
#| label: fig-FearNeighborhood
#| fig-cap: Bar graph showing the percentage of respondants, based on race and neighborhood diversity, that responded affirmative to GSS question, Is there any area right around here--that is, within a mile--where you would be afraid to walk alone at night?

gss_subset %>% 
  filter(!is.na(racecombo)) %>% 
  filter(racecombo != "Other") %>% 
  filter(!is.na(raclive)) %>%  
  group_by(racecombo, raclive) %>% 
  summarize(fear = sum(fear == "yes"), 
            n = n(),
            prop = fear/n) %>% 
  ggplot(aes(x = racecombo, y = prop)) +
  geom_col(aes(fill = raclive), position = "dodge") + 
  scale_fill_brewer(palette = "Set2") +
  scale_y_continuous(labels = scales::percent) +
  labs(title = "Social Fear by Race and Neighborhood Diversity",
       subtitle = "Is social fear influenced by respondent's race and their neighborhood's diversity?",
       x = "Race",
       y = "Percentage of Respondents who Expressed Fear") +
  theme_bw() +
  guides(fill = guide_legend(title = "Residential Diversity")) 
```

Lastly, we looked at sexual activity as a marker of social fear as represented in @fig-FearPartnersGender.

```{r}
#| label: fig-FearPartnersGender
#| fig-cap: Bar graph showing the percentage of respondants, by gender and number of sexual partners in the past 5 years, that responded affirmative to GSS question, Is there any area right around here--that is, within a mile--where you would be afraid to walk alone at night?

gss_subset %>% 
  filter(!is.na(partnrs5)) %>% 
  filter(!is.na(sexnow1)) %>% 
  filter(sexnow1 != "None of these") %>% 
  group_by(partnrs5, sexnow1) %>% 
  summarize(fear = sum(fear == "yes"), 
            n = n(),
            prop = fear/n) %>% 
  ggplot(aes(x = partnrs5, y = prop)) +
  geom_col(aes(fill = sexnow1), position = "dodge") + 
  scale_fill_brewer(palette = "Spectral") +
  scale_y_continuous(labels = scales::percent) +
  labs(title = "Social Fear by Gender and Sexual Partners",
       subtitle = "Is social fear influenced by respondent's gender and number of sexual partners?",
       x = "Number of Sexual Partners in the Past 5 Years",
       y = "Percentage of Respondents who Expressed Fear") +
  theme_bw() +
  guides(fill = guide_legend(title = "Gender"))

```

# Discussion

\newpage

# References